---
title: "SDM_Assignment3_3"
author: "Sri Balaji Muruganandam"
date: "25/10/2021"
output: html_document
---

### Setting Working Directory
```{r}
rm(list = ls())
setwd("G:\\SDM_Sem01\\Assignment3")
```

### Importing necessary libraries
```{r}
library(ISLR)
library(caret)
library(class)
library(ggplot2)
library(e1071)
library(gridExtra)
library(mvtnorm)
library(MASS)
```

### Loading the diabetes dataset
```{r}
load("Diabetes.RData")
dim(Diabetes)
```

### Exploring the high level overview of the data
```{r}
head(Diabetes,5)
str(Diabetes)
summary(Diabetes)
```

### Plotting Scatter Plot
```{r}
plot(Diabetes)
```
### There is a positive correlation between Glufast and the GluTest
```{r}
plot(Diabetes$glufast, Diabetes$glutest,col="#00918E", xlab = "Glufast", ylab = "Glutest", ylim = c(200,1600), main = "Correlation of Glufast and the GluTest", sub="They have positive correlation", lwd = 2.3)
```
## Scatter plot of five variables - relwt, glufast, glutest, instest, sspg
```{r}
table(Diabetes$group)
#Diabetes$group = as.numeric(Diabetes$group)
```
### relwt
```{r}

d1 = qplot(Diabetes$relwt, Diabetes$glutest, colour = Diabetes$group ,data = Diabetes)
d2 = qplot(Diabetes$relwt, Diabetes$instest, colour = Diabetes$group ,data = Diabetes)
d3 = qplot(Diabetes$relwt, Diabetes$sspg, colour = Diabetes$group ,data = Diabetes)
d4 = qplot(Diabetes$relwt, Diabetes$glufast, colour = Diabetes$group ,data = Diabetes)

grid.arrange(d1, d2, d3, d4, nrow = 2, ncol=2)
```

### glufast
```{r}
d1 = qplot(Diabetes$glufast,Diabetes$glutest, colour = Diabetes$group ,data = Diabetes)
d2 = qplot(Diabetes$glufast,Diabetes$instest, colour = Diabetes$group ,data = Diabetes)
d3 = qplot(Diabetes$glufast,Diabetes$sspg, colour = Diabetes$group ,data = Diabetes)

grid.arrange(d1, d2, d3, nrow = 2, ncol=2)
```

### glutest
```{r}
d1 = qplot(Diabetes$glutest,Diabetes$instest, colour = Diabetes$group ,data = Diabetes)
d2 = qplot(Diabetes$glutest,Diabetes$sspg, colour = Diabetes$group ,data = Diabetes)

grid.arrange(d1, d2, nrow = 1, ncol=2)
```

### Covariance 
```{r}
data2 = Diabetes
Diabetes$group = as.numeric(Diabetes$group)
covariance = cov(Diabetes)
print(covariance)
```
### Classes have strong correlation with the predictor glutest and sspg

### calculating multivariate normal
```{r}
norm_d = dmvnorm(x=Diabetes$group)
print(norm_d)
```



## (b)  Apply  linear  discriminant  analysis  (LDA)  and  quadratic  discriminant  analysis (QDA).    How  does  the  performance  of  QDA  compare  to  that  of  LDA  in  this  case?

### Preparing data
```{r}
#data2 = Diabetes
#set.seed(123)
#set.seed(121)
set.seed(123)
random_index = sample(c(1:nrow(Diabetes)), size = round(8/10 * nrow(Diabetes)), replace = FALSE)
train_data2 <- data2[random_index,]
test_data2 <- data2[-random_index,]

train_data2 = data.frame(train_data2)
test_data2 = data.frame(test_data2)
data2 = data.frame(data2)


y_train_data2 <- as.numeric(train_data2$group)-1
y_test_data2 <- as.numeric(test_data2$group)-1

#y_train_data2 <- train_data2$group
#y_test_data2 <- test_data2$group
```


## Modelling LDA
```{r}
lda_model = lda(group~., data = train_data2)
summary(lda_model)
```

### Predicting for test data
```{r}
lda_predict_train = predict(lda_model, newdata = train_data2)
lda_predict_test = predict(lda_model, newdata = test_data2)
res_train = as.numeric(lda_predict_train$class)-1
res_test = as.numeric(lda_predict_test$class)-1
#res_train = lda_predict_train$class
#res_test = lda_predict_test$class
```
### Calculating the train and test errors
```{r}

lda_result_train = which(res_train==y_train_data2)
lda_train_error=length(lda_result_train) / length(y_train_data2)
print(lda_train_error)


lda_result_test = which(res_test==y_test_data2)
lda_test_error=length(lda_result_test) / length(y_test_data2)
print(lda_test_error)
```

### The accuracy for training set is 90.51% and the accuracy for testing set is 86.20% when predicted using LDA.

## Modelling QDA
```{r}
qda_model = qda(group~., data = train_data2)
summary(qda_model)
```

### Predicting for test data
```{r}
qda_predict_train = predict(qda_model, newdata = train_data2)
qda_predict_test = predict(qda_model, newdata = test_data2)
res_trainq = as.numeric(qda_predict_train$class)-1
res_testq = as.numeric(qda_predict_test$class)-1
#res_trainq = qda_predict_train$class
#res_testq = qda_predict_test$class
```
### Calculating the train and test errors
```{r}

qda_result_train = which(res_trainq==y_train_data2)
qda_train_error=length(qda_result_train) / length(y_train_data2)
print(qda_train_error)


qda_result_test = which(res_testq==y_test_data2)
qda_test_error=length(qda_result_test) / length(y_test_data2)
print(qda_test_error)
```

### The accuracy for training set is 96.55% and the accuracy for testing set is 89.65% when predicted using QDA.

## QDA provides better prediction than the LDA

## (c)  Suppose  an  individual  has  (glucose  test/intolerence=  68,  insulin  test=122, SSPG =  544.  Relative  weight  =  1.86,  fasting  plasma  glucose  =  184).    To  which class  does  LDA  assign  this  individual?    To  which  class  does  QDA?

```{r}
#to_predict = c(1.86, 184, 68, 122, 544)
new_data1 <-data.frame(1.86, 184, 68, 122, 544)

colnames(new_data1)<-c("relwt","glufast","glutest","instest","sspg")
head(new_data1)
new_data_predict = predict(qda_model, newdata = new_data1)
print(new_data_predict$class)
```

## The class is found to be Overt_Diabetic for the given predictor values when predicted using QDA model

```{r}
new_data_predict_lda = predict(lda_model, newdata = new_data1)
print(new_data_predict_lda$class)
```

## The class is found to be Normal for the given predictor values when predicted using LDA model


