---
title: "SDM_Assignment5_5"
author: "Sri Balaji Muruganandam"
date: "8/12/2021"
output: html_document
---

### Setting Working Directory
```{r}
rm(list = ls())
setwd("G:\\SDM_Sem01\\Assignment5")
```

### Importing necessary libraries
```{r}
options(warn=-1)
library(rpart)
library(gbm)
library(randomForest)
library(geneplotter)
options(warn=0)
```

### Loading the Spam dataset
```{r}
load("spam.RData")
dim(spam)
```

### Exploring the high level overview of the data
```{r}
names(spam)
head(spam,5)
```

### Splitting the data into training and the test data
```{r}
set.seed(23)
random_index = sample(c(1:nrow(spam)), size = round(8/10 * nrow(spam)), replace = FALSE)
train_data <- spam[random_index,]
test_data <- spam[-random_index,]
#train_data = data.frame(train_data)
#test_data = data.frame(test_data)
data1 = data.frame(spam)

y_train_data <- as.numeric(train_data$spam)-1
y_test_data <- as.numeric(test_data$spam)-1
dim(train_data)
dim(test_data)
```

### Getting a Single Tree
```{r}
model_control = rpart.control(minsplit = 8, xval = 10, cp = 0)
model_fit = rpart(spam~., data = train_data, method = "class", control = model_control)
```

### Plotting the tree
```{r}
plot(model_fit)
text(model_fit, use.n = TRUE, cex = 0.3)
```

### Pruning the tree
```{r}
minimum_cp = which.min(model_fit$cptable[,4])
plot(model_fit$cptable[,4], xlab = "CP", ylab = "CV Error" )
```

### Prune Fit
```{r}
pruned_model_fit = prune(model_fit, cp = model_fit$cptable[minimum_cp, 1])
plot(pruned_model_fit)
text(pruned_model_fit, use.n = TRUE, cex = .5)
```

### Predicting for new values - For Test Data
```{r}
model_predict = predict(pruned_model_fit, newdata = test_data, type = "class")
y_predict = as.numeric(model_predict)-1
```

### Calculating Error
```{r}
prediction_error = sum(abs(y_predict - y_test_data))/length(y_test_data)
print(prediction_error)
```

## Modelling Random Forest for the spam data
### For Random Forest the Y column should be a factor. So Converting the Spam column to factor again
```{r}
train_data$spam = as.factor(train_data$spam)
model_fit_random_forest = randomForest(spam~., data = train_data, n.tree = 1000)
```

### Plotting the model parameters
```{r}
names(model_fit_random_forest)
varImpPlot(model_fit_random_forest)
importance(model_fit_random_forest)
```

### Predicting for new set of values - Test data
```{r}
y_predict = predict(model_fit_random_forest, newdata = test_data, type = "response")
y_predict = as.numeric(y_predict)-1
```

### Calculating Prediction error for Random Forest
```{r}
prediction_error_random_forest = sum(abs(y_predict - y_test_data))/length(y_test_data)
print(prediction_error_random_forest)
```

## Using Bagging for m = 10
```{r}
model_bagging <- randomForest(spam~., data = train_data, n.tree = 1000, mtry = 10)

varImpPlot(model_bagging)
importance(model_bagging)
```

### Predicting with bagging model
```{r}
y_predict = predict(model_bagging, newdata = test_data, type = "response")
y_predict = as.numeric(y_predict)-1
```

### Prediction Error for Bagging model
```{r}
prediction_error_bagging = sum(abs(y_predict - y_test_data))/length(y_test_data)
print(prediction_error_bagging)
```

## Modelling Random Forest for different values of m
### m = 3,4,5,6,7,8,9,10
```{r}
m = c(3,4,5,6,7,8,9,10)
error_rate = c()
for (i in 1:8){
model_bag_rf <- randomForest(spam~., data = train_data, n.tree = 350, mtry = m[i])

y_predict = predict(model_bag_rf, newdata = test_data, type = "response")
y_predict = as.numeric(y_predict)-1

predict_err_rf = sum(abs(y_predict - y_test_data))/length(y_test_data)
error_rate[i] = predict_err_rf
if(m[i]==3){
  m3 = model_bag_rf$err.rate[,1]
}
if(m[i]==5){
  m5 = model_bag_rf$err.rate[,1]
}
if(m[i]==7){
  m7 = model_bag_rf$err.rate[,1]
}
if(m[i]==9){
  m9 = model_bag_rf$err.rate[,1]
}
}
print(error_rate)
```

### Plotting m and the  test error value 
```{r}
par(bg = '#EEEEEE')
plot(m,error_rate, col="#E05D5D", type = "b", xlab = "m Value", ylab = " Test Error Rate", ylim = c(0,0.04), main = "m vs Test Error Rate", sub="Error is low when m = 5,6,7", lwd = 3.0)
```

### Plotting OOB Error rate and m
### Plotting for m values 3,5,7,9
```{r}
x_axis = c(1:length(m3))
plot(x_axis, m3, col = "#EC255A", type = "s", xlab = "No of trees given to the model", ylab
= "OOB Error Rate")
lines(x_axis, m5, col = "#161853", type = "s")
lines(x_axis, m7, col = "#064635", type = "s")
lines(x_axis, m9, col = "#DD4A48", type = "s")
```

## The Error is low when the value of m = 5,6,7

