---
title: "SDM_Assignment5_3"
author: "Sri Balaji Muruganandam"
date: "9/12/2021"
output: html_document
---

### Setting Working Directory
```{r}
rm(list = ls())
setwd("G:\\SDM_Sem01\\Assignment5")
```

### Importing necessary libraries
```{r}
options(warn=-1)
require(ISLR2)
require(caret)
require(e1071)
require(MASS)
options(warn=0)
```

### Loading the dataset
```{r}
data("OJ")
dim(OJ)
```

### Exploring the high level overview of the data
```{r}
names(OJ)
head(OJ,5)
```

### (a) Create a training set containing a random sample of 800 observations, and a test set containing the remaining observations.
### Splitting the data into training and the test data
```{r}
set.seed(23)
random_index = sample(c(1:nrow(OJ)), size = round(8/10 * nrow(OJ)), replace = FALSE)
train_data = OJ[random_index,]
test_data = OJ[-random_index,]

y_train_data = as.numeric(OJ$Purchase)-1
y_test_data = as.numeric(OJ$Purchase)-1
dim(train_data)
dim(test_data)
```

### (b) Fit a support vector classifier to the training data using cost = 0.01, with Purchase as the response and the other variables as predictors. Use the summary() function to produce summary statistics, and describe the results obtained.
```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'linear', cost = 0.01)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### (c) What are the training and test error rates?
### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### The training error is 0.5158879 and the test error is 0.5785047

## (d) Use the tune() function to select an optimal cost. Consider values in the range 0.01 to 10.
```{r}
model_tuning <- tune(svm, Purchase~., data = train_data, kernel = "linear", ranges = list(0.01, 0.1, 0.5, 1, 3.5, 5.5, 7.5, 10))
```

### Summary of tuned model
```{r}
summary(model_tuning)
names(model_tuning)
```

### Best value for cost
```{r}
print(model_tuning$best.parameters)
print(model_tuning$best.performance)
print(model_tuning$best.model)
```

### Best Value for optimal cost is 0.1612449

## (e) Compute the training and test error rates using this new value for cost.
### Computing with the cost of 0.1612449 and predicting the test and the train error

```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'linear', cost = 0.1612449)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### For Linear Kernel
### The training error is 0.5130841 and the test error is 0.5831776

## (f) Repeat parts (b) through (e) using a support vector machine with a radial kernel. Use the default value for gamma.

```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'radial', cost = 0.01)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### The training error is 0.6102804 and the test error is 0.6102804

```{r}
model_tuning <- tune(svm, Purchase~., data = train_data, kernel = "radial", ranges = list(0.01, 0.1, 0.5, 1, 3.5, 5.5, 7.5, 10))
```

### Summary of tuned model
```{r}
summary(model_tuning)
names(model_tuning)
```

### Best value for cost
```{r}
print(model_tuning$best.parameters)
print(model_tuning$best.performance)
print(model_tuning$best.model)
```

### Best Value for optimal cost is 0.1624624

### Computing with the cost of 0.1624624 and predicting the test and the train error

```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'radial', cost = 0.1624624)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### For Radical Kernel
### The training error is 0.5130841 and the test error is 0.5841121

## (g) Repeat parts (b) through (e) using a support vector machine with a polynomial kernel. Set degree = 2.

```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'polynomial', degree = 2, cost = 0.01)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### The training error is 0.6102804 and the test error is 0.611215

```{r}
model_tuning <- tune(svm, Purchase~., data = train_data, kernel = "polynomial",degree = 2, ranges = list(0.01, 0.1, 0.5, 1, 3.5, 5.5, 7.5, 10))
```

### Summary of tuned model
```{r}
summary(model_tuning)
names(model_tuning)
```

### Best value for cost
```{r}
print(model_tuning$best.parameters)
print(model_tuning$best.performance)
print(model_tuning$best.model)
```

### Best Value for optimal cost is 0.1928591

### Computing with the cost of 0.1928591 and predicting the test and the train error

```{r}
set.seed(23)
model_fit <- svm(Purchase~., data = train_data, kernel = 'polynomial', degree = 2, cost = 0.1928591)
```

### Getting the summary of the model
```{r}
summary(model_fit)
names(model_fit)
```

### Calculating for train data
```{r}
y_predict_train <- predict(model_fit, newdata = train_data, type = "response")
```
### Predicting Test Error
```{r}
y_predict_train = as.numeric(y_predict_train)-1
prediction_error_test = length(which(y_predict_train == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### Predicting for the new set of values - Test Data
```{r}
y_predict_test <- predict(model_fit, newdata = test_data, type = "response")
#print(model_predict)
```

### Calculating Prediction Error
```{r}
y_predict_test = as.numeric(y_predict_test)-1
prediction_error_test = length(which(y_predict_test == y_test_data))/length(y_test_data)
print(prediction_error_test)
```

### For Polinomial Kernel
### The training error is 0.5401869 and the test error is 0.5915888

## h) Overall, which approach seems to give the best results on this data?
### In the above three models, SVM with Linear Kernel provides better prediction than the other two models. Linear model provides train error of 5130841 and a test error of 0.5831776 which is less than the other models.