---
title: "SDM Assignment 1"
author: "Sri Balaji Muruganandam"
date: "17/09/2021"
output: html_document
---

### Setting Working Directory
```{r}
rm(list = ls())
setwd("G:\\SDM_Sem01\\Assignment1")
```

### Importing necessary libraries
```{r}
library(skimr)
library(dplyr)
```

### Importing Cereal data from local path
```{r}
c_data <- read.delim("cereal.csv",sep = ",")
dim(c_data)
```

### Viewing the Sample data
```{r}
head(c_data, 5)
tail(c_data, 5)
```

### Getting High Level Overview of the data
```{r}
str(c_data)
summary(c_data)
```

### Checking if there are any missing values
```{r}
sum(is.na(c_data))
```
There is no missing data in the dataset

### Outliers and Graphs
```{r}
boxplot(c_data$potass, notch = TRUE, main = "Potassium in Milligrams")
```
##### There are outliers in the potass column

### Removing the Outliers
##### As the data is very less. Keeping all of the data for Regression model.
```{r}
#boxplot.stats(c_data$potass)$out
#out_potass <- boxplot.stats(c_data$potass)$out
#c_data <- filter(c_data, potass != out_potass)
#dim(c_data)
#head(c_data, 10)
```

### Calories is normally distributed
```{r}
hist(c_data$calories,breaks = 10, col="#4B3869", border = "#6D8299", main = "Calories - Normally Distributed")

```

### There is a positive correlation between fiber and potass
```{r}
pairs(c_data$fiber~c_data$potass)
```



### Checking the unique values of mfr and type
```{r}
table(c_data$mfr)
table(c_data$type)
```

### Converting categorical variables into numerical values
```{r}
mfr_fact <- as.factor(c_data$mfr)
c_data$mfr <- as.numeric(mfr_fact)

c_data$type<-ifelse(c_data$type=="C",1,0)
```

## Data Transformation
### Converting vitamins(percentage) to float value

```{r}
table(c_data$vitamins)
c_data$vitamins <- c_data$vitamins / 100

#c_data$vitamins_num <- log(c_data$vitamins, base = 10)/2
#c_data$vitamins_num<-ifelse(c_data$vitamins_num =="-Inf",0,c_data$vitamins_num)
```

## Converting rating to float value
```{r}
c_data$rating <- c_data$rating / 100
```


### Taking log to Potass Column
#### Values of the Potass varying so much.
```{r}
table(c_data$potass)
c_data$potass <- log(c_data$potass, base = 10)
c_data <- na.omit(c_data)

head(c_data, 10)
sum(is.na(c_data))
```

### Saving the preprocessed data in Rdata file
```{r}
save(c_data, file = "cereal_clean_data.RData")
```



